---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
dataset <- read.csv('activity.csv')
head(dataset[is.na(dataset$steps)==FALSE & dataset$steps > 0,], n=10)
typeof(dataset)
typeof(dataset$steps)
typeof(dataset$date)
typeof(dataset$interval)
```

## What is mean total number of steps taken per day?
Calculate the total number of steps taken per day.
```{r}
# Include the plyr library to be able to use the ddply function
library(plyr)
# Summarize the dataset by date
daily <- ddply(dataset, "date", summarize, total.steps = sum(steps))
head(daily, n=10)
```

Make a histogram of the total number of steps taken each day.
```{r}
hist(daily$total.steps, breaks=20)
```

Calculate and report the mean and median of the total number of steps taken per day.
```{r}
daily.mean <- mean(daily$total.steps, na.rm=TRUE)
daily.median <- median(daily$total.steps, na.rm=TRUE)
```
The mean daily steps taken is `r sprintf("%.2f", daily.mean)` and the median daily steps taken is `r daily.median`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
